% =====================================================================
%  Physics and Code for Beginners -- A Guide to the Lunar-HFE Repository
%  Self-contained teaching handbook. Compile with:
%      latexmk -pdf -interaction=nonstopmode beginners_guide.tex
%  Designed to compile out-of-the-box on Overleaf (no shell-escape).
% =====================================================================
\documentclass[11pt,oneside]{article}

% ---- page geometry: generous, textbook-like margins -----------------
\usepackage[a4paper,margin=1in,headheight=14pt]{geometry}

% ---- fonts & encoding ----------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}

% ---- math ----------------------------------------------------------
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\sisetup{per-mode=symbol, inter-unit-product=\ensuremath{\cdot}}

% ---- tables, graphics, lists ---------------------------------------
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{enumitem}
\usepackage{caption}
\captionsetup{font=small,labelfont=bf,labelsep=period}

% ---- colour palette (matches the generated figures) ----------------
\usepackage{xcolor}
\definecolor{coral}{HTML}{B85B3A}
\definecolor{teal}{HTML}{2A6478}
\definecolor{forest}{HTML}{3D6E4A}
\definecolor{plum}{HTML}{5A4A6A}
\definecolor{char}{HTML}{2A2520}
\definecolor{dim}{HTML}{6E6862}
\definecolor{paper}{HTML}{FAF7F2}
\definecolor{softgrey}{HTML}{E8E5E0}
\definecolor{codebg}{HTML}{F6F3EE}

% ---- callout boxes --------------------------------------------------
\usepackage[most]{tcolorbox}
\tcbuselibrary{skins,breakable}

\newtcolorbox{plainenglish}[1][]{%
  enhanced, breakable, colback=teal!6, colframe=teal,
  boxrule=0.6pt, arc=2pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries, coltitle=teal,
  title={In plain English}, #1}

\newtcolorbox{analogy}[1][]{%
  enhanced, breakable, colback=forest!6, colframe=forest,
  boxrule=0.6pt, arc=2pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries, coltitle=forest,
  title={Analogy}, #1}

\newtcolorbox{keyidea}[1][]{%
  enhanced, breakable, colback=coral!7, colframe=coral,
  boxrule=0.8pt, arc=2pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries, coltitle=coral,
  title={Key idea}, #1}

\newtcolorbox{tryit}[1][]{%
  enhanced, breakable, colback=plum!6, colframe=plum,
  boxrule=0.6pt, arc=2pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries, coltitle=plum,
  title={Try it yourself}, #1}

\newtcolorbox{headsup}[1][]{%
  enhanced, breakable, colback=dim!8, colframe=dim,
  boxrule=0.6pt, arc=2pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries, coltitle=dim,
  title={Heads-up}, #1}

% ---- code listings (no shell-escape needed) ------------------------
\usepackage{listings}
\lstdefinestyle{pystyle}{%
  language=Python,
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\footnotesize\color{char},
  keywordstyle=\color{teal}\bfseries,
  commentstyle=\color{forest}\itshape,
  stringstyle=\color{coral},
  numberstyle=\tiny\color{dim},
  numbers=left, numbersep=8pt,
  showstringspaces=false,
  breaklines=true, breakatwhitespace=true,
  frame=single, rulecolor=\color{softgrey},
  framesep=6pt, xleftmargin=14pt, xrightmargin=4pt,
  tabsize=4, columns=fullflexible, keepspaces=true,
  literate={~}{{\textasciitilde}}1
}
\lstdefinestyle{shellstyle}{%
  language=bash,
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\footnotesize\color{char},
  keywordstyle=\color{teal}\bfseries,
  commentstyle=\color{dim}\itshape,
  showstringspaces=false,
  breaklines=true,
  frame=single, rulecolor=\color{softgrey},
  framesep=6pt, xleftmargin=8pt, xrightmargin=4pt,
  columns=fullflexible, keepspaces=true,
  literate={~}{{\textasciitilde}}1
}
\lstset{style=pystyle}

% ---- TikZ -----------------------------------------------------------
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,calc,fit,backgrounds,decorations.pathreplacing}

% ---- hyperref (load near last) -------------------------------------
\usepackage{hyperref}
\hypersetup{%
  colorlinks=true, linkcolor=teal, urlcolor=coral, citecolor=plum,
  pdftitle={Physics and Code for Beginners: A Guide to the Lunar-HFE Repository},
  pdfauthor={Lunar-HFE project}, bookmarksnumbered=true}

% ---- handy macros ---------------------------------------------------
\newcommand{\Kd}{\ensuremath{K_d}}
\newcommand{\Kdstar}{\ensuremath{K_d^{*}}}
\newcommand{\Ks}{\ensuremath{K_s}}
\newcommand{\Qb}{\ensuremath{Q_b}}
\newcommand{\code}[1]{\texttt{\small #1}}
\newcommand{\file}[1]{\texttt{\small #1}}
\newcommand{\mwmk}{\unit{mW.m^{-1}.K^{-1}}}

% =====================================================================
\begin{document}

% ---------- TITLE PAGE ----------
\begin{titlepage}
\centering
\vspace*{2.2cm}
{\color{dim}\large A hands-on companion to the Lunar-HFE research code}\\[1.6cm]
{\color{char}\Huge\bfseries Physics and Code\\[0.35cm] for Beginners}\\[0.8cm]
{\color{coral}\LARGE Measuring how the Moon's soil conducts heat}\\[2.0cm]

\begin{tcolorbox}[enhanced, width=0.82\textwidth, colback=paper,
  colframe=teal, boxrule=0.8pt, arc=3pt, halign=flush left]
\textbf{Who this is for.} You --- someone who is smart and motivated but has
\emph{no prior background} in coding, numerical methods, or heat-transfer
physics, and who wants to understand \emph{this specific project} from the
ground up.\\[4pt]
\textbf{The whole project in one sentence.} We take the only two sets of
underground temperature measurements ever made on the Moon (Apollo~15 and~17),
build a small physics simulator of how heat moves through Moon soil, and tune
one number --- the deep thermal conductivity \Kd{} --- until the simulator
matches the real measurements at each site.
\end{tcolorbox}

\vfill
{\color{dim}\small Generated teaching handbook \,\textbullet\, build the PDF with
\texttt{latexmk -pdf beginners\_guide.tex}}\\[2pt]
{\color{dim}\small Headline result: \Kdstar{}~$=$~\num{4.58}~\mwmk{} (Apollo~15),
\num{8.12}~\mwmk{} (Apollo~17).}
\vspace*{1.2cm}
\end{titlepage}

% ---------- TABLE OF CONTENTS ----------
\tableofcontents
\listoffigures
\clearpage

% ---------- HOW TO READ ----------
\section*{How to read this guide}
\addcontentsline{toc}{section}{How to read this guide}
Read it top to bottom, once, slowly. Each section builds on the one before.
Wherever you see a file path like \file{lunar/solver.py} or a function name
like \code{solve\_periodic\_equilibrium}, that is a \emph{real thing} in the
repository you can open and look at. The guide is the map; the code is the
territory. The coloured boxes are signposts:

\begin{itemize}[leftmargin=1.4em,itemsep=2pt]
  \item \textcolor{teal}{\textbf{In plain English}} --- the jargon-free version.
  \item \textcolor{forest}{\textbf{Analogy}} --- an everyday picture to lean on.
  \item \textcolor{coral}{\textbf{Key idea}} --- the one thing to remember.
  \item \textcolor{plum}{\textbf{Try it yourself}} --- a small action at the keyboard.
  \item \textcolor{dim}{\textbf{Heads-up}} --- a subtlety or common confusion.
\end{itemize}
\clearpage

% ---------- SECTIONS ----------
\input{sections/01_bigpicture}
\input{sections/02a_notation}
\input{sections/02_physics}
\input{sections/03_numerics}
\input{sections/04_anchor}
\input{sections/05_retrieval}
\input{sections/06_codetour}
\input{sections/07_running}
\input{sections/08_glossary}
\input{sections/09_roadmap}

\end{document}
